---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setoptions}
library(knitr)
opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data
```{r preprocessing}
activities <- read.csv("activity.csv")
head(activities)
```

## What is mean total number of steps taken per day?
```{r meanTotalSteps}
date <- factor(activities$date)
stepsByDate <- split(activities$steps, date)
sumStepsByDate <- as.numeric(lapply(stepsByDate, sum, na.rm = TRUE))
mean <- mean(sumStepsByDate)
med <- median(sumStepsByDate)
hist(sumStepsByDate, col = "green", xlab = "total number of steps", 
     main = "Histogram of total number of steps")
```

The mean of total number of steps is `r mean`.

The medium of total number of steps is `r med`.

## What is the average daily activity pattern?

```{r averageDailyActivity}
interval <- factor(activities$interval)
stepsByInterval <- split(activities$steps, interval)
meanStepsByInterval <- as.numeric(lapply(stepsByInterval, mean, na.rm = TRUE))
maxInterval <- which.max(meanStepsByInterval)
plot(meanStepsByInterval, type = "l", xlab = "Interval index", ylab = "Mean steps")
```

* The 5-minute interval, which contains the maximum number of steps is `r maxInterval`.

## Imputing missing values

```{r imputingMissingValues}
activities <- read.csv("activity.csv")
# Get the total number of NAs in the dataset
numOfNAs <- sum(is.na(activities$steps))
# Filling in missing values
# Replace the NA with the mean of that day
date <- factor(activities$date)
stepsByDate <- split(activities$steps, date)
meanStepsByDate <- as.numeric(lapply(stepsByDate, mean, na.rm = TRUE))
meanStepsByDate[is.nan(meanStepsByDate)] <- 0
newSteps <- c()
for (i in 1:length(stepsByDate)) {
    s <- as.vector(stepsByDate[[i]])
    s[is.na(s)]  <- meanStepsByDate[i]
    newSteps <- c(newSteps, s)
}

activities <- transform(activities, steps = newSteps)
stepsByDate <- split(activities$steps, date)
sumStepsByDate <- as.numeric(lapply(stepsByDate, sum))
mean <- mean(sumStepsByDate)
med <- median(sumStepsByDate)
hist(sumStepsByDate, col = "green", xlab = "total number of steps", 
     main = "Histogram of total number of steps after imputing")
```

The total number of missing values in the dataset is `r numOfNAs`

The mean of total number of steps after imputing missing values is `r mean`.

The medium of total number of steps after imputing missing values is `r med`.

These two values are exactly the same as part 1. My conclusion is that if we use
the mean for that day to replace the missing value of that day, there is no impact of
imputing missing data.

## Are there differences in activity patterns between weekdays and weekends?
```{r weekday}
date <- activities$date
weekday <- weekdays(as.Date(date))
weekday[weekday == "Saturday" | weekday == "Sunday"] <- "weekend"
weekday[weekday != "weekend"] <- "weekday"
activities <- transform(activities, weekday = weekday)
weekdayActivities <- activities[activities$weekday == "weekday",]
weekendActivities <- activities[activities$weekday == "weekend",]

interval <- factor(weekdayActivities$interval)
weekdayStepsByInterval <- split(weekdayActivities$steps, interval)
weekdayMeanStepsByInterval <- as.numeric(lapply(weekdayStepsByInterval, mean))

interval <- factor(weekendActivities$interval)
weekendStepsByInterval <- split(weekendActivities$steps, interval)
weekendMeanStepsByInterval <- as.numeric(lapply(weekendStepsByInterval, mean))

interval <- unique(activities$interval)
weekday <- rep("weekend", length(weekendMeanStepsByInterval))
weekendMeanActivities <- data.frame(steps = weekendMeanStepsByInterval,
                                    interval = interval,
                                    weekday = weekday)

weekday <- rep("weekday", length(weekdayMeanStepsByInterval))
weekdayMeanActivities <- data.frame(steps = weekdayMeanStepsByInterval,
                                    interval = interval,
                                    weekday =  weekday)

meanActivities <- rbind(weekendMeanActivities, weekdayMeanActivities)

library(lattice)
with(
  meanActivities,
  xyplot(steps ~ interval | weekday, type = "l", layout=c(1,2),
         xlab = "5 min interval", main = "Comparison of Activity on Weekdays vs Weekends")
)
```

My conclusion is that there is difference between weekday and weekend activities.
It seems that weekdays are more active.